---
title: "Untitled"
author: "Jonathan Bourne"
date: "28/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

#  Setup


```{r, echo = FALSE}

packages <- c("readxl", "readr", "tidyverse", "lubridate", "ggcorrplot", "Rtsne", "zoo")

new.packages <- packages[!(packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

sapply(packages, library, character.only = TRUE)

library(ukcovid19)

base_folder <- file.path("/home/jonno", "COVID_project")
data_folder <- file.path(base_folder, "COVID_project_data")

#list.files(data_folder)
```

create some useful functions

```{r, results='hide',echo = FALSE}
create_cor <- function(covid_msoa_time, target_var, use = "everything", method ="kendall"){
  
  cor_mat <-covid_msoa_time %>%
select(msoa.code, date, {{target_var}}) %>%
  pivot_wider(., values_from = {{target_var}}, names_from = msoa.code) %>%
  select(-date) %>%
  cor(., use = use, method = method)


percentiles_df <- as_tibble(cor_mat) %>% mutate(name1 = rownames(cor_mat)) %>% 
  pivot_longer(cols = -name1, names_to = "name2") %>%
  filter(name1!=name2) %>%
  mutate(percentile = percent_rank(value))
  
list(cor_mat = cor_mat, percentiles = percentiles_df)
}

```


load lsoa msoa lookup data

```{r, results='hide',echo = FALSE}
geog_lookup <- read_csv(file.path(data_folder, "Output_Area_to_LSOA_to_MSOA_to_Local_Authority_District_(December_2017)_Lookup_with_Area_Classifications_in_Great_Britain.csv")) %>%
  filter(RGN11NM == "London") %>%
  rename(msoa.code = MSOA11CD)
{
msoa_class <- geog_lookup %>%
  group_by(SOAC11CD, SOAC11NM, msoa.code) %>%
  summarise(counts = n()) %>%
  group_by(msoa.code) %>%
  mutate(fract = counts/sum(counts)) %>%
  arrange(-fract) %>%
  slice(1) %>%
  ungroup %>%
  mutate(class2 = ifelse(fract<.5, "mixed type", SOAC11NM)) 

msoa_class <- msoa_class %>%
  left_join(., msoa_class %>%
  group_by(class2) %>%
  summarise(class_counts = n())) %>%
  mutate(class3 = ifelse(class_counts <30, "other", class2))

}

#uses the CRDC loac classifications for London residence types
{
loac_df <- read_csv(file.path(data_folder, "LOAC_Lookup.csv")) %>%
  left_join(tibble(supgrp_cd = LETTERS[1:8], supgrp_name = c("Intermediate lifestyles", 
                                                 "High-density and high-rise flats",
                                                 "Settled Asians",
                                                 "Urban elites",
                                                 "City vibe",
                                                 "London life-cycle",
                                                 "Multi-ethnic suburbs",
                                                 "Ageing city fringe"))
)

  msoa_loac <- geog_lookup %>%
    left_join(loac_df, by = c("OA11CD"= "oa_code")) %>%
  group_by(supgrp_cd, supgrp_name, msoa.code) %>%
  summarise(counts = n()) %>%
  group_by(msoa.code) %>%
  mutate(fract = counts/sum(counts)) %>%
  arrange(-fract) %>%
  slice(1) %>%
  ungroup %>%
  mutate(class2 = ifelse(fract<.5, "mixed type", supgrp_name)) 

msoa_loac <- msoa_loac %>%
  left_join(., msoa_loac %>%
  group_by(class2) %>%
  summarise(class_counts = n())) %>%
  mutate(class3 = ifelse(class_counts <30, "other", class2))
  
}

#table(msoa_loac$class3)
#table(msoa_class$SOAC11NM)
#table(msoa_class$class2)
#table(msoa_class$class3)

```
load deprivation data

```{r, results='hide',echo = FALSE}
deprivation_df <- read_excel(file.path(data_folder, "File_1_-_IMD2019_Index_of_Multiple_Deprivation.xlsx"),  
                             sheet = "IMD2019",
                   .name_repair = "universal")  %>%
  set_names(., nm = tolower(names(.))) %>%
  select(lsoa.code = "lsoa.code..2011.", imd_rank = "index.of.multiple.deprivation..imd..rank")


```


load population estimates

```{r, results='hide',echo = FALSE}
#This chunk is used to load the population data for normalising the case numbers to per 100k population


#load data and make lower case headings with spaces as periods
pop_estimates_df <- read_excel(file.path(path = data_folder,"SAPE22DT4-mid-2019-msoa-syoa-estimates-unformatted.xlsx"), 
                   sheet = "Mid-2019 Persons",
                   skip = 3,
                   .name_repair = "universal")  %>%
  set_names(., nm = tolower(names(.)))



```


## Load the MSOA case count data

This chunk loads the data and ensures that all msoa have all time periods. empty time periods are assumed to have 0 new cases in that week.

It also adds the population estimates from 2019 to make case rates by 100k persons

A new variable is created which is the case data detrended

```{r, results='hide',echo = FALSE}

{
covid_msoa_time <- read_csv(file.path(data_folder, "covid_uk_msoa_07202021.csv")) %>%
  select(-X1) %>%
  arrange(date) %>%
  rename(rolling_sum = newCasesBySpecimenDateRollingSum,
         msoa.code = areaCode) %>%
  group_by(msoa.code) %>%
  mutate(current_cases =lag(rolling_sum) ) %>%
  ungroup %>%
  filter(date> ymd("2020-05-01"))

covid_msoa_time <- expand_grid(date = unique(covid_msoa_time$date), msoa.code = unique(covid_msoa_time$msoa.code)) %>%
  left_join(covid_msoa_time) %>%
  mutate(rolling_sum = ifelse(is.na(rolling_sum), 0, rolling_sum))%>%
  #add in population data
  left_join(pop_estimates_df %>% select(msoa.code, pop= all.ages)) %>%
  mutate(case_rate = (rolling_sum/pop)*1e5) %>%
  group_by(date) %>%
  mutate(rolling_centered = rolling_sum-mean(rolling_sum),
         case_rate_centered = case_rate-mean(case_rate)) %>%
  ungroup 
}


test <- covid_msoa_time %>%
  filter(!complete.cases(.))

```

Whilst there is a clear relationship between case_rate and rolling sum, it is heteroscedastic. 
```{r}
covid_msoa_time %>%
  ggplot(aes(x = rolling_sum, y = case_rate)) + geom_point()

```

# check case rate timeseries

```{r}


covid_msoa_time %>%
  pivot_longer(cols = c(case_rate, case_rate_centered)) %>%
  ggplot(aes(x = date, y = value, group = msoa.code)) +geom_line(alpha = 0.01) +
  facet_wrap(~name)

```
# Distributions

Centering the case rate about the mean for that week across all msoa removes the skew, showing that the data is normally distributed about the weekly mean

```{r}

covid_msoa_time %>%
  pivot_longer(cols = c(case_rate, case_rate_centered)) %>%
  #  filter(name != "case_rate") %>%
  ggplot(aes(x = value)) +geom_density() +
    facet_wrap(~name,scales = "free_x")

```


# Explore correlations

As case rate is skewed pearsons cannot be used, kendall is used instead.

Looking at the results shows that the centered data has much lower correlation and also a larger spread.

```{r}

case_rate_cor <- create_cor(covid_msoa_time = covid_msoa_time, target_var = case_rate,  method = "kendall")
case_rate_centered_cor <- create_cor(covid_msoa_time = covid_msoa_time, target_var = case_rate_centered,  method = "kendall")
case_rate_centered_cor_pearson <- create_cor(covid_msoa_time = covid_msoa_time, target_var = case_rate_centered,  method = "pearson")


list(case_rate_cor, case_rate_centered_cor, case_rate_centered_cor_pearson) %>%
  map_df(~{
    .x$percentiles %>%
  filter(percentile>0.025 & percentile<0.975) %>%
  arrange(value) %>%
  slice(1, n())
  }) %>%
  mutate(type = c("case_rate", "case_rate", "centered", "centered", "pearson", "pearson"))



```

# Visualising correlations

Althought the plots are not that great, they show that there are more obvious clusters in the centered data
```{r}
ggcorrplot(case_rate_cor$cor_mat, hc.order = TRUE) + theme_void()

```

```{r}
ggcorrplot(case_rate_centered_cor_pearson$cor_mat, hc.order = TRUE)  + theme_void()
```


# tSNE embeddings    

mapping resident classes and deprivation over a tsne embedding of the time series. reveals that there is a relationship with both. 

```{r}


  test <-covid_msoa_time %>%
  group_by(msoa.code) %>%
  #this little bit tries to leave only the signal ignoring the mean and the magnitude of the time series
  mutate(rolling_centered_mean  = mean(rolling_centered),
         rolling_centered_sd  = sd(rolling_centered),
         rolling_centered = (rolling_centered-rolling_centered_mean)/rolling_centered_sd) %>%
  ungroup %>%
  #little bit ends here
select(msoa.code, date, rolling_centered) %>%
  pivot_wider(., values_from = rolling_centered, names_from = date) 

set.seed(65)
test2 <- test %>%
  select(-msoa.code) %>%
  Rtsne(theta = 0.1)


test3 <- test %>% select(msoa.code) %>%
  mutate(d1 = test2$Y[,1],
         d2 = test2$Y[,2])

test3 %>%
  left_join(msoa_loac, by = "msoa.code") %>%
  filter(class3 !="mixed type") %>%
  ggplot(aes(x = d1, y = d2, colour = class3)) + geom_point() +
  scale_color_brewer(type = "qual")



 
covid_msoa_time %>%
  group_by(msoa.code) %>%
  #this little bit tries to leave only the signal ignoring the mean and the magnitude of the time series
  mutate(rolling_centered_mean  = mean(rolling_centered),
         rolling_centered_sd  = sd(rolling_centered),
         rolling_centered = (rolling_centered-rolling_centered_mean)/rolling_centered_sd) %>%
  ungroup %>%
  ggplot(aes(x = date, y = rolling_centered, group = msoa.code)) +geom_line(alpha = 0.01) 

# 
#looks like several normally distributed densities, the different groups might be the observed corellation clusters, can check this 
# covid_msoa_time %>%
#   group_by(msoa.code) %>%
#   #this little bit tries to leave only the signal ignoring the mean and the magnitude of the time series
#   mutate(rolling_centered_mean  = mean(rolling_centered),
#          rolling_centered_sd  = sd(rolling_centered),
#          rolling_centered = (rolling_centered-rolling_centered_mean)/rolling_centered_sd) %>%
#   ungroup %>%
#   ggplot(aes(x = rolling_centered, group = msoa.code), alpha = 0.01) + geom_density()

```

```{r}
dep2 <- deprivation_df %>%
  left_join(  geog_lookup %>%
  select(lsoa.code = LSOA11CD, msoa.code) %>%
    distinct()) %>%
  group_by(msoa.code) %>%
  summarise(imd_rank = mean(imd_rank)) %>%
  mutate(imd_rank_msoa = rank(imd_rank))

test3 %>%
  left_join(dep2) %>%
  ggplot(aes(x = d1, y = d2, colour = imd_rank_msoa)) + geom_point() +
  scale_color_viridis_c()

```


more deprived areas have higher total case rates of Covid

```{r}

all_time_covid <-covid_msoa_time %>%
  group_by(msoa.code) %>%
  summarise(rolling_sum = sum(rolling_sum),
            pop = first(pop)) %>%
  mutate(total_case_rate = rolling_sum/pop) %>%
  left_join(dep2) %>%
  mutate(percentile = percent_rank(imd_rank),
         target = case_when(
           percentile<0.5~"most deprived",
           percentile>0.5~"least deprived",
           TRUE ~"middle"
         ))
  

# 
# all_time_covid %>%
#   ggplot(aes(x = total_case_rate, colour = target)) + geom_density()

all_time_covid %>%
  ggplot(aes(x = imd_rank, y = total_case_rate, colour = target)) + geom_point()


all_time_covid %>%
  filter(target != "middle") %>%
  ggplot(aes(x = target, y = total_case_rate, fill = target)) + geom_boxplot()




t.test(all_time_covid$total_case_rate[all_time_covid$target=="most deprived"], all_time_covid$total_case_rate[all_time_covid$target=="least deprived"])

```

```{r}
all_time_covid %>%
  filter(target != "middle") %>%
  ggplot(aes(x = target, y = total_case_rate, fill = target)) + geom_boxplot()

```

# Cross-corellation

This chunk explores the lagged corellation between the 

```{r}

cor_mat <-covid_msoa_time %>%
select(msoa.code, date, case_rate_centered) %>%
  pivot_wider(., values_from = case_rate_centered, names_from = msoa.code) %>%
  select(-date)


test <-ccf(cor_mat[,1], cor_mat[,2], lag.max = 6)

test <-acf(cor_mat[,1:2], lag.max = 6)



test$acf

```



```{r}


new_tax_df <- read_excel(file.path(data_folder, "world_points_of_interest_premium_v6_6_0_Taxonomy.xlsx")) %>%
  rename(new_tax = ...9) %>%
  group_by(new_tax) %>%
  summarise(counts = n()) %>%
  arrange(-counts)

new_tax_df

```


#check mobility data

```{r}
test <- read_csv(file.path(data_folder, "data" ,"signals_set_2.csv"))
```

# hospitalisation  vs case lag
```{r}

cases_nation <- read_csv(file.path(data_folder, "cases_data_2021-Aug-03.csv")) %>%
  mutate(mean_7 = rollmean(newCasesBySpecimenDate, k = 7, fill = NA),
         type = "cases")

hospital_nation <- read_csv(file.path(data_folder, "hospital_data_2021-Aug-03.csv")) %>%
  mutate(mean_7 = rollmean(newAdmissions, k = 7, fill = NA),
         type = "hospital")

cases_nation %>%
  bind_rows(hospital_nation) %>%
  group_by(type) %>%
  mutate(mean = (mean_7 -mean(mean_7, na.rm = T))/sd(mean_7, na.rm = T)) %>%
  ungroup %>%
  ggplot(aes(x = date, y = mean, colour = type)) + geom_line() +
  labs(title = "Peak pattern between hospitalisations and cases", y = "rolling 7 day mean or normalised data")

```

